---
title: "Alexa_MLL_Biosynthesis_07172024"
output: pdf_document
date: "2024-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set working drive

```{r}
setwd("C:/Users/marqu/OneDrive - University of Denver/Desktop/DU Research/06272024_Alexa_Mextorquens_all_sample_batch_pos_10min/box plots/blank_removal_features")
```

## Load in neccesary packages

```{r}
library(ggplot2)
library(stringr)
library(data.table)
library(readr)
library(dplyr)
library(tidyr)
library(reshape)     
library(gplots)
library(pheatmap)
library(heatmap3)
library(magrittr) 
library(tibble) 
library(RColorBrewer)
library(paletteer)
library(clr)
library(tidyverse) #used for data science. The eight core packages inside this library are: ggplot2 (data visualisation), dplyr (data manipulation), tidyr, readr, purrr, tibble, stringr, and forcats
library(KODAMA) # to use the normalisation function
library(ggrepel) #mainly used to repel overlapping text labels in ggplots
library(vegan) #popular library for analysing ecological diversity and for multivariate analysis of community data. Here, we use it for PCoA
library(svglite) #to save the plots in support vector graphics (svg) format
library(factoextra) #for extracting and visualizing outputs of multivariate analyses such as PCA, k-means
library(ggsci) #provides color palettes for ggplot2 that can be used for scientific journals
library(matrixStats) #contains highly optimized functions to perform statistics on matrix data
library(cowplot) #efficient functions to arrange several plots
```

## Load in Data

### Normalized quant file

```{r}
quant <- read.csv("new_ft_Alexa_MLL_Mextorquens_07172024_with_blank_removal.csv", header=T, check.names=F)

```

### Metadata

```{r}
md <- read.csv("metadata_06272024_Alexa_Mextorquens.csv", header = T, check.names = F, sep = ',')
```

## Cleanup code

### Transpose the quant table

```{r}
quant <- t(quant)
quant <- as.data.frame(quant)

new_colnames <- as.character(quant[1, ])
quant <- quant[-1,]

colnames(quant) <- new_colnames

```

```{r}
numeric_columns <- sapply(quant, is.numeric)
numeric_columns

convert_to_numeric <- function(column) {
  as.numeric(as.character(column))
}

# Apply the function to all columns
quant <- quant %>% mutate_all(convert_to_numeric)
```

```{r}
quant <- quant %>% tibble::rownames_to_column("filename")

```

### Merge feature table and metadata for plotting

```{r}
data_merge <- md %>% dplyr::select("filename", "ATTRIBUTE_Sample_Name") %>% 
  left_join(quant, by = "filename")

data_merge <- md %>% dplyr::select("filename", "ATTRIBUTE_Metal") %>% 
  left_join(data_merge, by = "filename")
```

### Subset media blank and extraction blank

```{r}
data_merge <- data_merge %>% filter(ATTRIBUTE_Sample_Name != "Extraction_Blank")
data_merge <- data_merge %>% filter(ATTRIBUTE_Sample_Name != "Media_Blank")

```

### Set level to plot

```{r}
data_merge$ATTRIBUTE_Sample_Name <- factor(data_merge$ATTRIBUTE_Sample_Name, levels = c("mxaF_NdCl3", "mxaF_4132_kan_NdCL3","mxaF_4133_kan_NdCL3","mxaF4134_NdCl3","mxaF_4135_NdCl3","mxaF_4136_NdCl3",
"mxaF_4138_NdCl3","mxaF_4132_kan_Nd2O3", 
                                                                                      "mxaF_4133_kan_Nd2O3", "mxaF4134_Nd2O3", 
                                                                                      "mxaF4135_Nd2O3", "mxaF4136_Nd2O3", "mxaF_4138_clean_Nd2O3",
                                                                                      "WT_4132_kan_Nd2O3", "WT_4133_kan_Nd2O3"))
```

```{r}
new_labels <- c("mxaF_NdCl3" = "mxaF",
                 "mxaF_4132_kan_NdCL3" = "mxaF/4132",
                 "mxaF_4133_kan_NdCL3" = "mxaF/4133",
                 "mxaF4134_NdCl3" = "mxaF/4134",
                 "mxaF_4135_NdCl3" = "mxaF/4135",
                 "mxaF_4136_NdCl3" = "mxaF/4136",
                 "mxaF_4138_NdCl3" = "mxaF/4138",
                 "mxaF_4132_kan_Nd2O3" = "mxaF/4132",
                 "mxaF_4133_kan_Nd2O3" = "mxaF/4133",
                 "mxaF4134_Nd2O3" = "mxaF/4134",
                 "mxaF4135_Nd2O3" = "mxaF/4135",
                 "mxaF4136_Nd2O3" = "mxaF/4136",
                 "mxaF_4138_clean_Nd2O3" = "mxaF/4138",
                 "WT_4132_kan_Nd2O3" = "WT/4132",
                 "WT_4133_kan_Nd2O3" = "WT/4132")
```

## Plotting Methylolanthanin and analog plots with blank removal data

### Methylolanthanin (+1) charge

```{r}
# Plot deferoxamine by ammonium acetate
MLL_1_scatter_plot <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= MLL_28198_799.424_4.249)) +
  geom_line(linetype = "dashed") +
   scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Normalized Peak Area (AU)", x= expression(paste("")))+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"))

MLL_1_scatter_plot

ggsave(file="MLL_Charge_1_scatter_plot.svg", plot=MLL_1_scatter_plot, width =9, height = 7 )

```

### Assymetric MLL (+1) charge

```{r}
Assym_MLL_1_scatter_plot <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= Assym_25841_815.419_3.964)) +
  geom_line(linetype = "dashed") +
   scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Normalized Peak Area (AU)", x= expression(paste("")))+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"))

Assym_MLL_1_scatter_plot

ggsave(file="Assym_MLL_Charge_1_scatter_plot.svg", plot=Assym_MLL_1_scatter_plot, width =9, height = 7 )
```

### Rhodopetrobactin (+1) charge

```{r}
Rhodo_1_scatter_plot <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= Rh_24556_831.414_3.799)) +
  geom_line(linetype = "dashed") +
   scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Normalized Peak Area (AU)", x= expression(paste("")))+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"))

Rhodo_1_scatter_plot

ggsave(file="Rhodo_Charge_1_scatter_plot.svg", plot=Rhodo_1_scatter_plot, width =9, height = 7 )
```

### Methylolanthanin (+2) charge

```{r}
MLL_2_intensity_scatter_plot <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= `28199_400.215_4.246`, shape= ATTRIBUTE_Metal)) +
  geom_line(linetype = "dashed") +
  scale_x_discrete(labels=new_labels)+
   scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Peak Intensity (AU)", x= expression(paste("")), shape ="")+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"),
        legend.text=element_text(size=14,family="sans"))

MLL_2_intensity_scatter_plot

ggsave(file="MLL_Charge_2_intensity_blank_removal_scatter_plot.svg", plot=MLL_2_intensity_scatter_plot, width =9, height = 7 )
```

### Assym. Methylolanthanin (+2) charge

```{r}
Assym_MLL_2_intensity_scatter_plot <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= `25829_408.213_3.951`, shape= ATTRIBUTE_Metal)) +
  geom_line(linetype = "dashed") +
  scale_x_discrete(labels=new_labels)+
   scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Peak Intensity (AU)", x= expression(paste("")), shape ="")+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"),
        legend.text=element_text(size=14,family="sans"))

Assym_MLL_2_intensity_scatter_plot

ggsave(file="Assym_MLL_Charge_2_intensity_blank_removal_scatter_plot.svg", plot=Assym_MLL_2_intensity_scatter_plot, width =9, height = 7 )
```

### Rhodopetrobactin (+2) charge

```{r}
Rhodo_2_intensity_scatter_plot <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= `24645_416.21_3.809`, shape= ATTRIBUTE_Metal)) +
  geom_line(linetype = "dashed") +
  scale_x_discrete(labels=new_labels)+
   scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Peak Intensity (AU)", x= expression(paste("")), shape ="")+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"),
        legend.text=element_text(size=14,family="sans"))

Rhodo_2_intensity_scatter_plot
ggsave(file="Rhodo_Charge_2_intensity_blank_removed_scatter_plot.svg", plot=Rhodo_2_intensity_scatter_plot, width =9, height = 7 )
```

## Does 4135 KO have an effect on acylation of MLL for compound 8 or 12

```{r}
feature_7202_559mz <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= `7202_559.381_1.416`, shape= ATTRIBUTE_Metal)) +
  geom_line(linetype = "dashed") +
  scale_x_discrete(labels=new_labels)+
   scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Peak Intensity (AU)", x= expression(paste("")), shape ="")+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"),
        legend.text=element_text(size=14,family="sans"))

feature_7202_559mz

ggsave(file="feature_7202_559_intensity_blank_removal_mz.svg", plot=feature_7202_559mz, width =9, height = 7 )
```

```{r}
feature_24132_559mz <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= f24132_559.381_3.753)) +
  geom_line(linetype = "dashed") +
  scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Normalized Peak Area (AU)", x= expression(paste("")))+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"))

feature_24132_559mz

ggsave(file="feature_24132_559mz.svg", plot=feature_24132_559mz, width =9, height = 7 )
```

## Compound 9 abundance across samples

```{r}
feature_23578_721mz <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= f23578_721.412_3.651)) +
  geom_line(linetype = "dashed") +
  scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Normalized Peak Area (AU)", x= expression(paste("")))+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"))

feature_23578_721mz

ggsave(file="feature_23578_721mz.svg", plot=feature_23578_721mz, width =9, height = 7 )
```

```{r}
feature_23756_361mz <- data_merge %>%
  tail(60) %>%
  ggplot( aes(x= ATTRIBUTE_Sample_Name, y= f23756_361.21_3.661)) +
  geom_line(linetype = "dashed") +
  scale_y_log10()+
  geom_point(size= 3) +
  theme_classic() + 
  labs(y= "Normalized Peak Area (AU)", x= expression(paste("")))+
  theme(axis.title.x = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.title.y = element_text(family = "sans", size = 16, hjust = 0.5),  
        axis.text.x = element_text(family = "sans", size = 16, color = "black", angle = 45, hjust =1), 
        axis.text.y = element_text(family = "sans", size = 16, color= "black"),
        axis.line=element_line(size =1),
        axis.ticks=element_line(size=1), 
        axis.ticks.length=unit(.25, "cm"))

feature_23756_361mz

ggsave(file="feature_23756_361mz.svg", plot=feature_23756_361mz, width =9, height = 7 )
```

## Compound 9 abundance across samples
